---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# tardis

<!-- badges: start -->
<!-- badges: end -->
*Functions for processing AG501 EMA data and converting to EMU-compatible formats*

The tardis package contains functions for

1. processing position-calculated and head-corrected EMA sensor data from the Carstens AG501 system, including calculating derivatives and further derived variables
2. converting processed files to the SSFF format for use with the EMU system
3. calculating MFCCs from WAV files and converting these to SSFF format.
4. wrapper functions for emuR, such as obtaining multiple trackdata objects in a single call

These functions can be used independently, although conversion to SSFF makes some assumptions about the structure of your AG501 .txt files. Please check the vignettes for extensive documentation and example usage.

This package is in-progress and currently undergoing testing. Use at your own caution!

## Installation

You can install tardis as follows:

```{r eval=FALSE}
remotes::install_github("samkirkham/tardis", build_vignettes = TRUE)
```

## Example

For documentation of typical workflow and example usage:

```{r eval=FALSE}
vignette("tardis-workflow", package = "tardis")
```

